---
title: "Case Study: Wealth and Life Expectancy"
author: "Ricardo Leon Coronado"
date: "3/3/2021"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r}

library(tidyverse)
library(gapminder)



# Basic plot with title
ggplot( data=gapminder, aes(x=lifeExp)) + 
  geom_histogram(fill="skyblue", alpha=0.5) +
  ggtitle("Life Expectancy by Country") +
  theme_minimal()
#p
```

DESCRIPTION: This graph takes all the values and its showing me what all life expectancies values amount to. We see from this graph that the higher proportion of life expectancy was around 70 to 72 yrs old.


```{r}

library(tidyverse)
library(gapminder)



# Basic plot with title
ggplot( data=gapminder, aes(x=continent)) + 
  geom_bar(fill="skyblue", alpha=0.5) +
  ggtitle("Continents") +
  theme_minimal()
#p
```

DESCRIPTION: This graph is another interesting one. Simple as it is, it divides all the counts by continent so I can see a more comprehensive view of every continent.

```{r}

library(tidyverse)
library(gapminder)



# Basic plot with title
ggplot( data=gapminder, aes(x=continent, y=lifeExp)) + 
  geom_boxplot(fill="skyblue", alpha=0.5) +
  ggtitle("Continents") +
  theme_minimal()
#p
```
DESCRIPTION: The last two graphs featured different types of values of data. The first one featured quantitative data, or basically, numbers, while the second one featured qualitative data, or in other words, categories. This third graph is a combination of both types. You have qualitative data and quantitative data combined in a single graph plot where they are all each divided by country.

```{r}

library(tidyverse)
library(gapminder)



#We apply a filter to get rid of Kuwait.
dat <- gapminder

view(dat)

dat1 <- dat %>% filter(country != "Kuwait")

view(dat1)
#p
```


```{r}
library(ggplot2)

ggplot(dat1, mapping = aes(x = lifeExp, y = gdpPercap, color = continent, size = pop/100000)) +
  geom_point() +
  facet_wrap(~ year, nrow = 1) +
  scale_y_continuous(trans = "sqrt")
  theme_bw()

ggsave("plot.png", width = 15, units = "in")
```