---
title: "Exploratory Data Analysis"
author: "Ricardo Leon Coronado"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Reading Notes

Replace this text with your reading notes.

## EDA Example

The code below is an example of the EDA process using the `starwars` data from the `tidyverse` package. (Make sure you have the `tidyverse` package installed!)

Run the code line-by-line and look at the output. Add a comment to each line of code that explains what it does/what insights it provides.

```{r}
#This loads the library that we want to use.
library(tidyverse)

#It gives us the dimensions of the dataframe, number of rows and number of columns.
dim(starwars)

#This gives us the name of the all the different columns in the database.
colnames(starwars)

#Head should give us the first few rows of the tables.
head(starwars)

#Glimpse should give us a more comprehensive view of the data that's in every column but not quite all of it,
glimpse(starwars)

#When you type a question mark before the name of a method or a dataset, you get more information or "help" on such data set or such method, and that shows up in your view window to the right.
?starwars

#The pipe operator is used to "pipe" in R. Upon research I learned that it basically takes the output of one function and passes it into another function as an argument.
starwars %>% count(species)

#This allows us to get the means of the values of one entire column, the one after the dollar sign. So, in other words, the "height" column.
mean(starwars$height)

#In this case, it's the na.rm what allows rows in the column that have NA to exist.
mean(starwars$height, na.rm = TRUE)

#The summary gives us the values of the Min, Median, Mean, Max, and the 1st and 3rd quartile altogether.
summary(starwars$height)

#This is a function to turn centimeters to feet and it's taking the measurement in centimeters and dividing it by 30.48 and returning the value and storing it inside the variable "ft".
cm_to_ft <- function(cm){
  ft = cm / 30.48
  return(ft)
}

#
starwars_2 <- starwars %>% mutate(height_ft = cm_to_ft(height)) 

#This will give us the dimensions of the new dataframe, starwars_2. Number of columns by number of rows.
dim(starwars_2)

#This will give us the all the different column names for the dataframe (starwars_2)
colnames(starwars_2)

#Similar to a previous summary, this command tell the computer that we want a summary of the values located specifically in the "height_ft" column in the dataframe.
summary(starwars_2$height_ft)

#This command produced a histogram with the numbers and values of the dataframe.
ggplot(starwars_2, aes(height_ft)) + 
  geom_histogram()

#I am not 100% sure on this one but I think it is giving us the value with the highest height in all our dataset.
starwars_2 %>% filter(height_ft == max(height_ft))
starwars_2 %>% filter(height_ft == max(height_ft, na.rm = TRUE))

# https://starwars.fandom.com/wiki/Yarael_Poof
```

## EDA Practice

Continue exploring the `starwars` data to gain additional insights, using [R4DS: Chapter 7](https://r4ds.had.co.nz/exploratory-data-analysis.html) as a guide.

It is ok if you don't understand the all code in Chapter 7. (That is what we'll be learning the next two weeks!) If writing your own code is a struggle, try the "copy, paste, and tweak" method.

```{r}
# your code goes here
```